Simulation study in rectangular domain

### 2024-04-26 R/simulation_study/simulation_study_fjords.Rmd
various updates, including removing parallelization in batches fit cause it takes too much RAM

```{r}
#number of batch
N_BATCH=1

@@

#duration of each track in hour
TMAX=0.5*24

@@

#get the coastline geometry from the geojson file
land<-st_read(file.path(greenland_data_path,"land_utm.shp"))
```

@@

hf=1/60 #1 minute time step

@@

#mixture of independent gaussian
ftau_gauss=function(cov_data,D0=0.5,tau0=1,tau1=3) {
  Dshore=cov_data$DistanceShore
  theta=cov_data$theta
  tau=tau0+(tau1-tau0)*exp(-((theta-pi/2)/(pi/5))^2-(Dshore/D0)^2)+(tau1-tau0)*exp(-((theta+pi/2)/(pi/5))^2-(Dshore/D0)^2)
  return (tau)
}

@@

ftau_step=function(cov_data,epsilon=pi/10,alpha=pi/5) {
    Dshore=cov_data$DistanceShore
    theta=cov_data$theta
    tau=1.5+0.5*(tanh((theta+(pi/2+epsilon))/alpha)-tanh((theta+(pi/2-epsilon))/alpha))+
      0.5*(tanh((theta+(-pi/2+epsilon))/alpha)-tanh((theta+(-pi/2-epsilon))/alpha))
    return (tau)
}

fomega=function(cov_data,D0=0.3,omega0=60*pi/2,lambda=2,kappa=0.2) {
    Dshore=cov_data$DistanceShore
    theta=cov_data$theta
    if (is.null(Dshore)){
      Dshore=1/cov_data$ExpShore
    }
    coeff=exp(-kappa*(Dshore/D0)^2)
    omega=omega0/2*(tanh(lambda*(theta+pi/2-atanh(-0.9)/lambda))+tanh(lambda*(theta-(pi/2-atanh(-0.9)/lambda))))*coeff
    return(omega)
}

@@

Dshore=seq(from=0.1,to=2,length.out=30)
theta=seq(from=-pi,to=pi,length.out=30)
grid<- as.data.frame(expand.grid(Dshore,theta))
colnames(grid)=c("DistanceShore","theta")
true_par=grid
true_par$tau=ftau_step(grid)
true_par$omega=fomega(grid)

plot_true_tau <- plot_ly(true_par,type = "mesh3d",
                         x = ~theta,y = ~DistanceShore,z=~tau,intensity=~tau,
                         colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("True tau"),scene=list(xaxis = list(title = "Theta",showgrid = F),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),zaxis = list(title = "tau")))

plot_true_omega <- plot_ly(true_par,type = "mesh3d",
                         x = ~theta,y = ~DistanceShore,z=~omega,intensity=~omega,
                         colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("True omega"),scene=list(xaxis = list(title = "Theta",showgrid = F),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),zaxis = list(title = "omega")))

plot_true_tau
plot_true_omega
```

@@

```{r}
#number of points
n <- 500

@@

@@

Dshore=seq(from=0.3,to=2,length.out=30)
theta=seq(from=-pi,to=pi,length.out=30)
grid<- as.data.frame(expand.grid(DistanceShore,theta))
colnames(grid)=c("DistanceShore","theta")
true_par=grid
true_par$omega=fomega_splines(grid)

plot_true_omega <- plot_ly(true_par,type = "mesh3d",
                         x = ~theta,y = ~DistanceShore,z=~omega,intensity=~omega,
                         colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("True omega"),scene=list(xaxis = list(title = "Theta",showgrid = F),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),
                                                            zaxis = list(title = "omega")))

plot_true_omega
```
## Simulation with low measurement errors and high frequency

@@

plot_samples_fjords_hf_ne=ggplot()+geom_sf(data=land$geometry,fill=NA)+coord_sf(datum=st_crs(32626))+
  geom_point(data=data_fjords_hf_ne$BATCH1,mapping=aes(y1,y2,color=ID),size=0.1)+
  geom_path(data=data_fjords_hf_ne$BATCH1,mapping=aes(y1,y2,color=ID),size=0.1)+
  geom_point(data = data_fjords_hf_ne$BATCH1%>% filter(!duplicated(ID)),
             aes(x = y1, y = y2), shape = 3, size = 4, col = "red")+
  xlab("x") + ylab("y")

>>>>

```{r}
#number of batch
N_BATCH=10

@@

#duration of each track in hour
TMAX=1*24

@@

#get the land and coastline geometries from the geojson file
land<-st_read(file.path(greenland_data_path,"land_utm.shp"))
coastline<-st_read(file.path(greenland_data_path,"coastline_utm.geojson"))
```

@@

hf=1/120 #1 minute time step

@@

#mixture of independent gaussian
ftau_gauss=function(cov_data,D0=1,tau0=1,tau1=3,alpha=pi/10) {
  Dshore=cov_data$DistanceShore
  theta=cov_data$theta
  tau=tau0+(tau1-tau0)*exp(-((theta-pi/2)/alpha)^2-(Dshore/D0)^2)+(tau1-tau0)*exp(-((theta+pi/2)/alpha)^2-(Dshore/D0)^2)
  return (tau)
}

@@

ftau_bump=function(cov_data,epsilon=pi/10,alpha=pi/5,D0=1) {
    Dshore=cov_data$DistanceShore
    theta=cov_data$theta
    tau=(1.5+0.5*(tanh((theta+(pi/2+epsilon))/alpha)-tanh((theta+(pi/2-epsilon))/alpha))+
      0.5*(tanh((theta+(-pi/2+epsilon))/alpha)-tanh((theta+(-pi/2-epsilon))/alpha)))
    return (tau)
}

fomega=function(cov_data,D0=0.1,omega0=60*pi,lambda=2,kappa=0.2) {
    Dshore=cov_data$DistanceShore
    theta=cov_data$theta
    if (is.null(Dshore)){
      Dshore=1/cov_data$ExpShore
    }
    coeff=exp(-kappa*(Dshore/D0)^2)
    omega=omega0/2*(tanh(lambda*(theta+pi/2-atanh(-0.9)/lambda))+tanh(lambda*(theta-(pi/2-atanh(-0.9)/lambda))))*coeff
    return(omega)
}

@@

Dshore=seq(from=0.1,to=2,length.out=30)
theta=seq(from=-pi,to=pi,length.out=30)
grid<- as.data.frame(expand.grid(Dshore,theta))
colnames(grid)=c("DistanceShore","theta")
true_tau=matrix(ftau_constant(grid),30,30)
true_omega=matrix(fomega(grid),30,30)

plot_true_tau <- plot_ly(type = "surface",
                         x = ~theta,y = ~Dshore,z=~true_tau,
                         colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("True tau"),
                       scene=list(xaxis = list(title = "theta",showgrid = F),
                      yaxis = list(title = "Distance to shore",showgrid = F),zaxis = list(title = "tau")))

plot_true_omega <- plot_ly(type = "surface",
                           contours = list(z = list(show = TRUE, start = min(true_omega), end = max(true_omega), size = 5,color="black")),
                         x = ~theta,y = ~Dshore,z=~true_omega,
                         colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("True omega"),
                       scene=list(xaxis = list(title = "Theta",showgrid = F),
                      yaxis = list(title = "Distance to shore",showgrid = F),zaxis = list(title = "omega")))

plot_true_tau
plot_true_omega

@@

```{r}
#number of points
n <- 1000

@@

#k-th basis spline function
fomega_spline_basis=function(cov_data,k=1) {

  if (is.null(cov_data$ExpShore)){
    Xp=predict(m1, newdata = cov_data, type = "lpmatrix")

  }
  else if (is.null(cov_data$DistanceShore)){
    Xp=predict(m2, newdata = cov_data, type = "lpmatrix")
  }
  coeffs=rep(0,ncol(Xp))
  coeffs[k]=1
  return(as.numeric(Xp%*%coeffs))
}

@@

Dshore=seq(from=0.3,to=2,length.out=30)
theta=seq(from=-pi,to=pi,length.out=30)
grid<- as.data.frame(expand.grid(Dshore,theta))
colnames(grid)=c("DistanceShore","theta")
spline_omega=matrix(fomega_splines(grid),30,30)

# Extract the basis function values for the k-th basis function on the grid
z2=matrix(fomega_spline_basis(grid,k=2),30,30)
z3=matrix(fomega_spline_basis(grid,k=3),30,30)
z4=matrix(fomega_spline_basis(grid,k=4),30,30)

plot_spline_omega <- plot_ly(type = "surface",
                             contours = list(z = list(show = TRUE, start = min(spline_omega), end = max(spline_omega), size = 5,color="black")),
                         x = ~Dshore,y = ~theta,z=~spline_omega,
                         colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("Spline smooth omega"),scene=list(xaxis = list(title = "Theta",showgrid = T),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),
                                                            zaxis = list(title = "omega")))

plot_basis2 <- plot_ly(type = "surface",contours = list(z = list(show = TRUE, start = min(z2), end = max(z2), size = 0.05,color="black")),
                         x = ~Dshore,y = ~theta,z=~z2, colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("Spline basis 2"),scene=list(xaxis = list(title = "Theta",showgrid = T),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),
                                                            zaxis = list(title = "omega")))

plot_basis3 <- plot_ly(type = "surface",contours = list(z = list(show = TRUE, start = min(z3), end = max(z3), size = 0.05,color="black")),
                         x = ~Dshore,y = ~theta,z=~z3, colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("Spline basis 3"),scene=list(xaxis = list(title = "Theta",showgrid = T),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),
                                                            zaxis = list(title = "omega")))

plot_basis4 <- plot_ly(type = "surface",contours = list(z = list(show = TRUE, start = min(z4), end = max(z4), size = 0.05,color="black")),
                         x = ~Dshore,y = ~theta,z=~z4, colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("Spline basis 4"),scene=list(xaxis = list(title = "Theta",showgrid = T),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),
                                                            zaxis = list(title = "omega"))) 
```

```{r}
plot_spline_omega
```

```{r}
plot_basis2
```

```{r}
plot_basis3
```

```{r}
plot_basis4
```

## Simulation with low measurement errors and high frequency

@@

plot_samples_fjords_hf_ne=ggplot()+geom_sf(data=land$geometry,fill=NA)+coord_sf(datum=st_crs(32626))+
  geom_point(data=data_fjords_hf_ne$BATCH1,mapping=aes(y1,y2,color=ID),size=0.1)+
  geom_path(data=data_fjords_hf_ne$BATCH1,mapping=aes(y1,y2,color=ID),size=0.1)+
  geom_point(data = data_fjords_hf_ne$BATCH2%>% filter(!duplicated(ID)),
             aes(x = y1, y = y2), shape = 3, size = 4, col = "red")+
  xlab("x") + ylab("y")

### 2024-04-26 R/simulation_study/simulation_study_rect.Rmd
change number of batches
various updates, including removing parallelization in batches fit cause it takes too much RAM

```{r global variables}
#number of batch
N_BATCH=10

#number of individual tracks per batch
N_ID=6

@@

#degree of freedom in tensor splines
SP_DF=5
```

@@

Dshore=seq(from=0.3,to=2,length.out=30)
theta=seq(from=-pi,to=pi,length.out=30)
grid<- as.data.frame(expand.grid(Dshore,theta))
colnames(grid)=c("DistanceShore","theta")
true_par=grid
true_par$tau=ftau_constant(grid)
true_par$omega=fomega(grid)

plot_true_tau <- plot_ly(true_par,type = "mesh3d",
                         x = ~theta,y = ~DistanceShore,z=~tau,intensity=~tau,
                         colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("True tau"),scene=list(xaxis = list(title = "Theta",showgrid = F),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),zaxis = list(title = "tau")))

plot_true_omega <- plot_ly(true_par,type = "mesh3d",
                         x = ~theta,y = ~DistanceShore,z=~omega,intensity=~omega,
                         colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("True omega"),scene=list(xaxis = list(title = "Theta",showgrid = F),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),zaxis = list(title = "omega")))

@@

### Plot

@@

Dshore=seq(from=0.3,to=2,length.out=30)
theta=seq(from=-pi,to=pi,length.out=30)
grid<- as.data.frame(expand.grid(DistanceShore,theta))
colnames(grid)=c("DistanceShore","theta")
true_par=grid
true_par$omega=fomega_splines(grid)

plot_true_omega <- plot_ly(true_par,type = "mesh3d",
                         x = ~theta,y = ~DistanceShore,z=~omega,intensity=~omega,
                         colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("True omega"),scene=list(xaxis = list(title = "Theta",showgrid = F),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),
                                                            zaxis = list(title = "omega")))

plot_true_omega
```

@@

crcvm_rect_hf_ne1<- foreach(i=1:N_BATCH,.packages=c("smoothSDE")) %dopar% {
  #estimation with tensor splines
  formulas <- list(mu1=~1,mu2=~1,tau=~s(ID,bs="re"),
                 nu=~s(ID,bs="re"),omega=~te(theta,ExpShore,k=SP_DF,bs="cs"))
  par0 <- c(0,0,1,1,0)
  m <- SDE$new(formulas = formulas,data = data_rect_hf_ne[[paste("BATCH",k,sep="")]],type = "RACVM",
                      response = c("y1","y2"),par0 = par0,fixpar=c("mu1","mu2"),
                      other_data=list("H"=H_hf))
  m$fit(method="BFGS")
  m 
}

@@

crcvm_rect_hf_ne2<- foreach(i=1:N_BATCH,.packages=c("smoothSDE")) %dopar% {
  #estimation with tensor splines
  formulas <- list(mu1=~1,mu2=~1,tau=~s(ID,bs="re"),
                 nu=~s(ID,bs="re"),omega=~te(theta,DistanceShore,k=SP_DF,bs="cs"))
  par0 <- c(0,0,1,1,0)
  m <- SDE$new(formulas = formulas,data = data_rect_hf_ne[[paste("BATCH",k,sep="")]],type = "RACVM",
                      response = c("y1","y2"),par0 = par0,fixpar=c("mu1","mu2"),
                      other_data=list("H"=H_hf))
  m$fit(method="BFGS")
  m 
}

>>>>

```{r global variables}
#number of batch
N_BATCH=1

#number of individual tracks per batch
N_ID=3

@@

#degree of freedom in tensor splines
SP_DF=5

@@

Dshore=seq(from=0.3,to=2,length.out=30)
theta=seq(from=-pi,to=pi,length.out=30)
grid<- as.data.frame(expand.grid(Dshore,theta))
colnames(grid)=c("DistanceShore","theta")
true_tau=matrix(ftau_constant(grid),30,30)
true_omega=matrix(fomega(grid),30,30)

plot_true_tau <- plot_ly(type = "surface",
                         x = ~Dshore,y = ~theta,z=~true_tau,
                         colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("True tau"),
                       scene=list(xaxis = list(title = "Theta",showgrid = F),
                      yaxis = list(title = "Distance to shore",showgrid = F),zaxis = list(title = "tau")))

plot_true_omega <- plot_ly(type = "surface",
                           contours = list(z = list(show = TRUE, start = min(true_omega), end = max(true_omega), size = 5,color="black")),
                         x = ~Dshore,y = ~theta,z=~true_omega,
                         colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("True omega"),
                       scene=list(xaxis = list(title = "Theta",showgrid = F),
                      yaxis = list(title = "Distance to shore",showgrid = F),zaxis = list(title = "omega")))

@@

### Plot spline omega and spline basis functions

@@

Dshore=seq(from=0.3,to=2,length.out=30)
theta=seq(from=-pi,to=pi,length.out=30)
grid<- as.data.frame(expand.grid(Dshore,theta))
colnames(grid)=c("DistanceShore","theta")
spline_omega=matrix(fomega_splines(grid),30,30)

# Extract the basis function values for the k-th basis function on the grid
z2=matrix(fomega_spline_basis(grid,k=2),30,30)
z3=matrix(fomega_spline_basis(grid,k=3),30,30)
z4=matrix(fomega_spline_basis(grid,k=4),30,30)

plot_spline_omega <- plot_ly(type = "surface",
                             contours = list(z = list(show = TRUE, start = min(spline_omega), end = max(spline_omega), size = 5,color="black")),
                         x = ~Dshore,y = ~theta,z=~spline_omega,
                         colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("Spline smooth omega"),scene=list(xaxis = list(title = "Theta",showgrid = T),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),
                                                            zaxis = list(title = "omega")))

plot_basis2 <- plot_ly(type = "surface",contours = list(z = list(show = TRUE, start = min(z2), end = max(z2), size = 0.05,color="black")),
                         x = ~Dshore,y = ~theta,z=~z2, colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("Spline basis 2"),scene=list(xaxis = list(title = "Theta",showgrid = T),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),
                                                            zaxis = list(title = "omega")))

plot_basis3 <- plot_ly(type = "surface",contours = list(z = list(show = TRUE, start = min(z3), end = max(z3), size = 0.05,color="black")),
                         x = ~Dshore,y = ~theta,z=~z3, colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("Spline basis 3"),scene=list(xaxis = list(title = "Theta",showgrid = T),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),
                                                            zaxis = list(title = "omega")))

plot_basis4 <- plot_ly(type = "surface",contours = list(z = list(show = TRUE, start = min(z4), end = max(z4), size = 0.05,color="black")),
                         x = ~Dshore,y = ~theta,z=~z4, colors=colorRamp(c("blue", "lightblue", "chartreuse3", "yellow", "red")))%>%
                layout(title=paste("Spline basis 4"),scene=list(xaxis = list(title = "Theta",showgrid = T),
                                                                 yaxis = list(title = "Distance to shore",showgrid = F),
                                                            zaxis = list(title = "omega"))) 
```

```{r}
plot_spline_omega
```

```{r}
plot_basis2
```

```{r}
plot_basis3
```

```{r}
plot_basis4
```

@@

crcvm_rect_hf_ne1<- foreach(i=1:N_BATCH,.packages=c("smoothSDE")) %do% {
  #estimation with tensor splines
  formulas <- list(mu1=~1,mu2=~1,tau=~s(ID,bs="re"),
                 nu=~s(ID,bs="re"),omega=~te(theta,ExpShore,k=SP_DF,bs="cs"))
  par0 <- c(0,0,1,1,0)
  m <- SDE$new(formulas = formulas,data = data_rect_hf_ne[[paste("BATCH",k,sep="")]],type = "RACVM",
                      response = c("y1","y2"),par0 = par0,fixpar=c("mu1","mu2"),
                      other_data=list("H"=H_hf))
  m$fit(method="BFGS")
  m 
}

@@

crcvm_rect_hf_ne2<- foreach(i=1:N_BATCH,.packages=c("smoothSDE")) %do% {
  #estimation with tensor splines
  formulas <- list(mu1=~1,mu2=~1,tau=~s(ID,bs="re"),
                 nu=~s(ID,bs="re"),omega=~te(theta,DistanceShore,k=SP_DF,bs="cs"))
  par0 <- c(0,0,1,1,0)
  m <- SDE$new(formulas = formulas,data = data_rect_hf_ne[[paste("BATCH",k,sep="")]],type = "RACVM",
                      response = c("y1","y2"),par0 = par0,fixpar=c("mu1","mu2"),
                      other_data=list("H"=H_hf))
  m$fit(method="BFGS")
  m 
}
